---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pivotprobs

<!-- badges: start -->
<!-- badges: end -->

`pivotprobs` is a package for computing the probability of pivot events (election results where a single ballot can determine the winner) and other election events. 

It aims to offer both general methods that can handle any voting system and belief distribution and specialized methods that handle important cases efficiently.  



## Installation

You can install development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("aeggers/pivotprobs")
```

## Example

Lots of dependencies -- you also need `rcdd` and `geometry`. 

```{r example, message = F}
library(pivotprobs)
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)

alpha3 <- c(.4, .35, .25)*85
plurality3 <- plurality_election(k = 3)

sc_out <- plurality3 %>%
   election_event_probs(method = "sc", alpha = alpha3, tol = .1)
sc_out[["a_b"]]$integral
sc_out[["a_b"]]$seconds_elapsed
sc_out[["a_c"]]$integral
sc_out %>% map("seconds_elapsed") %>% unlist() %>% sum()

mc_out <- plurality3 %>%
   election_event_probs(method = "mc", alpha = alpha3, num_sims = 100000)
mc_out[["a_b"]]$integral
mc_out[["a_b"]]$seconds_elapsed
mc_out[["a_c"]]$integral
mc_out %>% map("seconds_elapsed") %>% unlist() %>% sum()

ev_out <- plurality3 %>%
  election_event_probs(method = "ev", alpha = alpha3)
ev_out[["a_b"]]$integral
ev_out[["a_b"]]$seconds_elapsed
ev_out[["a_c"]]$integral
ev_out %>% map("seconds_elapsed") %>% unlist() %>% sum()
```


Checking the Condorcet method 

